---
title: "wifi_fingerprinting"
author: "Stacey Jackson"
date: "13/05/2019"
output:
  html_document: default
  word_document: default
---

```{r}
if(require("pacman")=="FALSE"){
  install.packages('pacman')
  library('pacman')
  pacman::p_load(here, readxl, plyr, caret, dplyr, doParallel,
                 lubridate, corrplot, ggplot2, 
                 tidyverse, arules, arulesViz, rstudioapi,RMySQL,
                 plotly, padr, lubridate, forecast,zoo, htmltools, readr)
} else {
  library('pacman')
  pacman::p_load(here, readxl, plyr, caret, dplyr, doParallel,
                 lubridate, corrplot, ggplot2, 
                 tidyverse, arules, arulesViz, rstudioapi,RMySQL,
                 plotly, padr, lubridate, forecast,zoo,htmltools, readr)
}


#Upload the files
trainingData<-read.csv(file="/Users/staceyjackson/Dropbox (Personal)/Ubiqum/May_2019_Wifi/Dataset/UJIndoorLoc/trainingData.csv", header=TRUE, sep=",")

validationData<-read.csv(file="/Users/staceyjackson/Dropbox (Personal)/Ubiqum/May_2019_Wifi/Dataset/UJIndoorLoc/validationData.csv", header=TRUE, sep=",")
```

####1. INVESTIGATE THE TRAINING SET
####1.1 Factors and timestamp
```{r}

#Look at the classes
str(trainingData)
sapply(trainingData, class)


#change 523-528 to factors - relative position, USERID, PHONEID, BUILDING ID
columns <- c(523:528)
trainingData[,columns] <- lapply(trainingData[,columns], as.factor)

sapply(trainingData,class)


#change time to POSixt
trainingData$TIMESTAMP <- as.POSIXct(trainingData$TIMESTAMP,origin="1970-01-01")
validationData$TIMESTAMP <- as.POSIXct(validationData$TIMESTAMP,origin="1970-01-01")
```

#1.2. Initial plots
```{r}
plot(trainingData$BUILDINGID)
plot(trainingData$USERID)
plot(trainingData$SPACEID)

p <- ggplot(data = trainingData, mapping = aes(x =USERID, y=PHONEID))
p + geom_point()

p <- ggplot(data = validationData, mapping = aes(x =LONGITUDE, y=LATITUDE))
p + geom_point()

plot(trainingData$USERID)
plot(trainingData$PHONEID)

p <- ggplot(data=trainingData, mapping=aes(x=USERID, y=PHONEID))
p + geom_point()

p <- ggplot(data=trainingData, mapping=aes(x=USERID, y=BUILDINGID))
p + geom_point()

#plot of all three buildings' longitude & latitude showing each floor
P<-ggplot(trainingData, mapping=aes(x=LONGITUDE, y=LATITUDE, colour=FLOOR))
P+geom_point()


#create separate training datasets for each building and plot longitude & latitude showing each floor
trainingDataBuilding0<-trainingData%>%
  filter(BUILDINGID==0)

P<-ggplot(trainingDataBuilding0, mapping=aes(x=LONGITUDE, y=LATITUDE, colour=FLOOR))
P+geom_point()

trainingDataBuilding1<-trainingData%>%
  filter(BUILDINGID==1)
P<-ggplot(trainingDataBuilding1, mapping=aes(x=LONGITUDE, y=LATITUDE, colour=FLOOR))
P+geom_point()

trainingDataBuilding2<-trainingData%>%
  filter(BUILDINGID==2)
P<-ggplot(trainingDataBuilding2, mapping=aes(x=LONGITUDE, y=LATITUDE, colour=FLOOR))
P+geom_point()


```

#1.3. Delete WAPs with a mean of 100
```{r}

#means<-apply(trainingData[1:520], 2, mean)
#means<-as.data.frame(means)

mean<-0
for (i in 1:520){
  mean[i]<-mean(trainingData[,i])
}
mean[521:529]<-0
mean<-as.data.frame(mean)

#delete all the WAPs with a mean of =100
indices<-c()
indices[521:529]<-0
for (i in 1:520){
if (mean[i,]==100){
  indices[i]<- i
}
}
trainingData2<- trainingData[is.na(indices)]

trainingData2<-cbind(trainingData2,trainingData[,c(521:529)])
```
#1.4 Phones with weak signals
```{r}
phones <- split(trainingData2, trainingData2$PHONEID)

sapply(phones, function(x) {
         colMeans(x[, c(1:50)])
})

#variance of the phones
sapply(phones, var)
```
#1.5 WAPs with low variance
```{r}

variance <- sapply(trainingData2[,c(1:465)], var)
low_variance_WAP <- sapply(variance, function(x){
 (x>5)
})
low_variance_WAP<- which(low_variance_WAP, arr.ind = TRUE)
low_variance_WAP<- as.data.frame(low_variance_WAP)

indices2<-low_variance_WAP[,1]
trainingData3<- trainingData2[(indices2)]

trainingData4<-cbind(trainingData3,trainingData2[,c(466:474)])
```
#1.6 Change WAP values from 100 to -105 
```{r}

change_WAP_value <- apply(trainingData4[,c(1:430)], 2, function(x) {ifelse(x == 100, -105, x)})

change_WAP_value<-as.data.frame(change_WAP_value)

trainingData5<-cbind(change_WAP_value,trainingData2[,c(466:474)])

```

#1.7 Find WAP values >-30 and <0
```{r}
#found WAPs with high values but what to do with them?

high_value <- sapply(trainingData5[,c(1:430)], function(x){
 (x>=-30 & x<=0)
})

which(high_value, arr.ind = TRUE)
```

#1.8 Means and sds of the WAPs
```{r}


trainingData5_mean<-apply(trainingData5[,c(1:430)], 2, mean)
trainingData5_sd<-apply(trainingData5[,c(1:430)], 2, sd)

```

#2. REMOVE VARIABLES - spaceID, relative position, userID, phoneID, timestamp
```{r}
trainingData6<-trainingData5[,-c(435:439)]
```

#3. CREATE SAMPLE OF THE TRAINING SET
```{r}

sample <- trainingData6 %>% group_by(FLOOR, BUILDINGID) %>% sample_n(100)

table(sample$FLOOR)
table(sample$BUILDINGID)

#plot not working
a <- htmltools::tagList()    
for(i in unique(sample$BUILDINGID)){
a[[i]] <- sample %>% dplyr:: filter(BUILDINGID == i) %>% plot_ly(type = "scatter3d",
       x = ~ LATITUDE,
       y = ~ LONGITUDE,
       z = ~ FLOOR,
       mode = 'markers')


}  
a[[1]]
a[[2]]
a[[3]]
```
#4. MODEL 1: randomForest 
```{r}
# Load package
library(randomForest)

# Saving the waps in a vector
WAPs<-grep("WAP", names(trainingData6), value=T)

# Get the best mtry
bestmtry_rf<-tuneRF(sample[WAPs], sample$LONGITUDE, ntreeTry=100,stepFactor=2,improve=0.05,trace=TRUE, plot=T) 

# Train a random forest using that mtry
system.time(rf_reg<-randomForest(y=sample$LONGITUDE,x=sample[WAPs],importance=T,method="rf", ntree=100, mtry=22))

# Train a random forest using caret package
system.time(rf_reg_caret<-train(y=sample$LONGITUDE, x=sample[WAPs], data = sample, method="rf", ntree=100,tuneGrid=expand.grid(.mtry=22)))
```
```


